---
title: "cjoint model to make pairwise comparison statements"
format: html
---

0. this updates conjoint_design.R from NK
1. restricted-randomization conjoint design object with constraints:
* NEW ok for sender = recipient; it means that there is no further data sharing
* constraint 1 - health concern purpose blocked for non-clinical senders (now excludes head coach too)
* constraint 2 - load management blocked for clinical senders (head coach ALLOWED)
* constraint 3 - team readiness blocked for clinical senders (head coach ALLOWED)
* constraint 4 - health concern cannot route to sports scientist or whole team
* constraint 5 - load management cannot route to whole team
* NEW constraint 6 - sender = head coach, purpose = team readiness, recipient excludes all but head coach, whole team
* NEW constraint 7 - sender = head coach, purpose = health concern, recipient excludes all non-clinical including head coach

```{r}
#| echo: true
#| eval: true
#| message: false
#| warning: false
library(jsonlite)
library(cjoint)
library(dplyr)
library(tibble)

spec <- fromJSON("../../data/conjoint/attributes_levels.json")
attribute_levels <- list(
  sender    = spec$attributes$sender,
  purpose   = spec$attributes$purpose,
  recipient = spec$attributes$recipient
)

# ---- Constraints, expressed in cjoint's constraint-list format ----
# Each constraint lists levels for one attribute that are jointly infeasible
# with levels for another attribute. cjoint expects one constraint entry per
# blocked (attribute_A level combination) -> (attribute_B levels) pairing.
constraints <- list(
  list( # C1 health concern, no non-clinician sender
    purpose = "health concern",
    sender  = c("strength coach", "sports scientist")
  ),
  list( # C2 load management, no clinician sender
    purpose = "load management",
    sender  = c("physician", "athletic trainer")
  ),
  list( # C3 team readiness, no clinician sender
    purpose = "team readiness",
  #  sender  = c("physician", "athletic trainer", "head coach")
    sender  = c("physician", "athletic trainer")
  ),
  list( # C4 health concern, no sport scientist or whole team recipient
    purpose   = "health concern",
    recipient = c("sports scientist", "whole team")
  ),
  list( # C5 load management, no whole team recipient
    purpose   = "load management",
    recipient = "whole team"
  ),
  list( # C6 sender = head coach and purpose = team readiness, recipient excludes all but head coach, whole team
    sender  = "head coach",
    purpose = "team readiness",
    recipient = c("physician","athletic trainer","sports scientist","strength coach")
  ),
  list( # C7 sender = head coach, purpose = health concern, recipient excludes all non-clinical including head coach
    sender  = "head coach",
    purpose = "health concern",
    recipient = c("sports scientist","strength coach","whole team","head coach")
  )
)

# ---- Marginal probabilities of each level BEFORE restriction ----
# cjoint uses these as the baseline distribution to reweight from when
# excluding the constrained combinations above.
level.probs <- list(
  sender    = rep(1 / length(attribute_levels$sender), length(attribute_levels$sender)),
  purpose   = rep(1 / length(attribute_levels$purpose), length(attribute_levels$purpose)),
  recipient = rep(1 / length(attribute_levels$recipient), length(attribute_levels$recipient))
)
```
2. cjoint model object makes it easy to generate statements as strings

```{r}
design <- makeDesign(
  type = "constraints",
  attribute.levels = attribute_levels,
  constraints = constraints,
  level.probs = level.probs
)

saveRDS(design, "../../data/conjoint/cjoint_design.rds")
```

3. create pairwise comparison statements
* 39 statements (increase from 35 in NK)

```{r}
# dataframe with all attributes-levels combinations
indices <- which(design$J > 0, arr.ind = TRUE)
combinations <- apply(indices, 1, function(idx) {
  sapply(seq_along(idx), function(dim_i) {
    dimnames(design$J)[[dim_i]][[idx[dim_i]]]
  })
})
df <- as.data.frame(t(combinations), stringsAsFactors = FALSE)
colnames(df) <- names(dimnames(design$J))
rownames(df) <- paste0("V", sprintf("%03d", 1:nrow(df)))

# transform attributes-levels to survey statements for pairwise comparison
phrases <- c(
  "sender_readiness" = ", along with the rest of the team's data",
  "health concern" = " evaluates the data for a health concern",
  "load management" = " evaluates the data to decide whether to increase or reduce your training load that day",
  "team readiness" = " aggregates the data"
)

df <- df %>%
  mutate(value = paste0(
    "You send your menstrual cycle data to your team's ",
    sender,
    if_else(purpose == "team readiness", phrases["sender_readiness"], ""),
    ". The ",
    sender,
    phrases[purpose],
    if_else(sender == recipient, 
            " and does not share the information further.", 
            paste0(" and passes it along to your ", recipient,".")
    )
  ))
json_df <- df %>%
  rownames_to_column(var = "id")
write_json(json_df, "../../data/conjoint/conjoint_CI_cjoint.json", pretty = TRUE)
json_df
```
---
