---
title: "update to conjoint parameters, constraints, model"
format: html
---

1.  read, display CI parameters (sender, purpose, recipient) for conjoint analysis

```{r}
#| echo: true
#| eval: true
#| message: false
#| warning: false
library(jsonlite)

spec <- fromJSON("../../data/conjoint/attributes_levels.json")
spec

```

2.  set constraints

- problems arose with constraints that separated "purpose" into two distinct graphs
- distinct graphs results in parallel regression terms that fail to intersect
- added C6, C7 to satisfy network requirement & prevent parallel regression terms
- sender = recipient is OK; it just means that information is not shared

```{r}
#| echo: true
#| eval: true
#| message: false
#| warning: false
constraints <- list(
  # C1: health concern purpose blocked for non-clinical senders (now excludes head coach too)
  list(
    purpose = "health concern",
    sender  = c("strength coach", "sports scientist")
  ),
  # C2a: load management blocked for clinical senders 
  # ***head coach now ALLOWED here***
  # represents an athlete's direct, voluntary self-disclosure to her coach for a
  # practical training decision; this is the "ethical but plausible" pathway,
  # distinct from a third party forwarding clinical data without her involvement)
  list(
    purpose = "load management",
    sender  = c("physician", "athletic trainer")
  ),
  # C2b: team readiness blocked for clinical senders AND head coach (aggregation is
  # a performance-staff function only)
  # ***head coach not excluded***
  list(
    purpose = "team readiness",
  #  sender  = c("physician", "athletic trainer", "head coach")
    sender  = c("physician", "athletic trainer")
  ),
  # C4: health concern cannot route to sports scientist or whole team
  list(
    purpose   = "health concern",
    recipient = c("sports scientist", "whole team")
  ),
  # C5: load management cannot route to whole team
  list(
    purpose   = "load management",
    recipient = "whole team"
  ),
  # C6: sender = head coach and purpose = team readiness
  # recipient excludes all but head coach, whole team
  list(
    sender  = "head coach",
    purpose = "team readiness",
    recipient = c("physician","athletic trainer","sports scientist","strength coach")
  ),
  # C7: sender = head coach and purpose = health concern
  # recipient excludes all non-clinical including head coach
  list(
    sender  = "head coach",
    purpose = "health concern",
    recipient = c("sports scientist","strength coach","whole team","head coach")
  )
)


```

3.  create cjoint model

```{r}
#| echo: true
#| eval: true
#| message: false
#| warning: false
library(cjoint)

# ---- Attribute levels (must match survey_design.json / vignette_items.json) ----
attribute_levels <- list(
  sender    = spec$attributes$sender,
  purpose   = spec$attributes$purpose,
  recipient = spec$attributes$recipient
)

# ---- Marginal probabilities of each level BEFORE restriction ----
# cjoint uses these as the baseline distribution to reweight from when
# excluding the constrained combinations above.
level.probs <- list(
  sender    = rep(1 / length(attribute_levels$sender), length(attribute_levels$sender)),
  purpose   = rep(1 / length(attribute_levels$purpose), length(attribute_levels$purpose)),
  recipient = rep(1 / length(attribute_levels$recipient), length(attribute_levels$recipient))
)

design <- makeDesign(
  type = "constraints",
  attribute.levels = attribute_levels,
  constraints = constraints,
  level.probs = level.probs
)

saveRDS(design, "../../data/cjoint_design.rds")
```

4.  combinations of attributes and levels

```{r}
#| echo: true
#| eval: true
#| message: false
#| warning: false
library(dplyr)

# dataframe with all attributes-levels combinations
indices <- which(design$J > 0, arr.ind = TRUE)
combinations <- apply(indices, 1, function(idx) {
  sapply(seq_along(idx), function(dim_i) {
    dimnames(design$J)[[dim_i]][[idx[dim_i]]]
  })
})
df <- as.data.frame(t(combinations), stringsAsFactors = FALSE)
colnames(df) <- names(dimnames(design$J))
rownames(df) <- paste0("V", sprintf("%03d", 1:nrow(df)))

```

5.  produce, display statements list from model

```{r}
#| echo: true
#| eval: true
#| message: false
#| warning: false
library(tibble)

# transform attributes-levels to survey statements for pairwise comparison
phrases <- c(
  "sender_readiness" = ", along with the rest of the team's data",
  "health concern" = " evaluates the data for a health concern",
  "load management" = " evaluates the data to decide whether to increase or reduce your training load that day",
  "team readiness" = " aggregates the data"
)

df <- df %>%
  mutate(value = paste0(
    "You send your menstrual cycle data to your team's ",
    sender,
    if_else(purpose == "team readiness", phrases["sender_readiness"], ""),
    ". The ",
    sender,
    phrases[purpose],
    if_else(sender == recipient, 
            " and does not share the information further.", 
            paste0(" and passes it along to your ", recipient,".")
    )
  ))
json_df <- df %>%
  rownames_to_column(var = "id") %>%
  toJSON(auto_unbox = TRUE, pretty = TRUE)
write_json(json_df, "../../data/CI.json", auto_unbox = TRUE, pretty = TRUE)
json_df

```
